First steps towards a formalization of Forcing

### 2018-06-18 LSFA/rs.tex

A different version of $\DC$ (without the constraint $f(0)=a$, but
allowing a parameter that allows to prove an equivalence with $\AC$) was
formalized as part of the Isabelle/ZF library, along with many
statements that follow from or are equivalent to $\AC$. Since the
proofs there required a greater generality, it would have been very
contrived to derived this ``pointed'' version of $\DC$ from that
development. Instead, we preferred to give a direct proof based in the
one that appears in Moschovakis \cite{moschovakis1994notes}. Another
advantage is that this proof depends on the following version of $\AC$
(existence of choice functions on $\P(X)\sm \{\mty\}$):
\[
\exists s: \P(A)\sm \{\mty\}\to A \, \forall X\sbq A.\ X\neq \mty \limp
s(X) \in X.
\]
This statement is more standard than the one deemed ``$\AC$'' in the
Isabelle/ZF development, that involves a silent use of Replacement
since it is written as an axiom scheme.

@@

The function \verb|dc_witness| has, appart from $A$, $a$ and $R$, a
function $s$ as a parameter. If this function is a selector for
$\P(A)\sm \{\mty\}$, the function $f\defi$\verb|dc_witness|$(A,a,s,R)$
will satify $\DC$.

>>>>

A different version of $\DC$ (without the constraint $f(0)=a$, but
allowing a parameter that allows to prove an equivalence with $\AC$) was
formalized as part of the Isabelle/ZF library, along with many
statements that follow from or are equivalent to $\AC$. Since the
proofs there required a greater generality, it would have been very
contrived to derived this ``pointed'' version of $\DC$ from that
development. Instead, we preferred to give a direct proof based in the
one that appears in Moschovakis \cite{moschovakis1994notes}. Another
advantage is that this proof depends on the following version of $\AC$
(existence of choice functions on $\P(A)\sm \{\mty\}$):
\[
\exists s: \P(A)\sm \{\mty\}\to A \, \forall X\sbq A.\ X\neq \mty \limp
s(X) \in X.
\]
This statement is more standard than the one deemed ``$\AC$'' in the
Isabelle/ZF development, that involves a silent use of Replacement
since it is written as an axiom scheme.

@@

The function \verb|dc_witness| has, apart from $A$, $a$ and $R$, a
function $s$ as a parameter. If this function is a selector for
$\P(A)\sm \{\mty\}$, the function $f\defi$\verb|dc_witness|$(A,a,s,R)$
will satify $\DC$.
